---
title: "p8105_hw1_ksf2129"
author: Kimia Faroughi
output: github_document
date: "2025-09-16"
---

```{r}
library(moderndive)
library(tidyverse)
```

## Problem 1

Load the early_january_weather dataset.

```{r}
data("early_january_weather")
```

Look at the data.

```{r}
early_january_weather #view dataset
help(early_january_weather) #open dataset help file
skimr::skim(early_january_weather) #skim data
```

# Description of the Dataset

The dataset shows hourly meteorological data for LGA, JFK, and EWR airports for January 2013. 

Variables include: `r names(early_january_weather)`

Origin refers to the weather station/airport, month as 1 denoting January, hour starts with 0 and goes to 23, "temp" as temperature in Fahrenheit, "dewp" as dewpoint in Fahrenheit, "humid" as relative humidity, "wind_dir" as wind direction in degrees, "wind_speed" as wind speed in miles per hour, "wind_gust" as gust speed in miles per hour, "precip"  as precipitation in inches, "pressure" as sea level pressure in millibars, "visib" as visibility in miles, and "time_hour" as date and hour of the recording. 

There are `r nrow(early_january_weather)` rows (observations) and `r ncol(early_january_weather)` columns (variables) in the dataset.

The mean temperature (in Fahrenheit) is `r mean(early_january_weather$temp)`. 

# Scatterplot

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point() #generate scatterplot
ggsave("problem1_scatterplot.pdf", height = 4, width = 6) #save scatterplot
```

The scatterplot shows a cyclical temperature pattern for the first half of January. The temperature seems to rise and fall throughout the day each day. The maximum value these cycles reach is steadily increasing from the beginning to the middle of the month, with the minimum temperatures steadily rising as well. Additionally, the humidity is generally low for a majority of the time period, but around January 10th, the humidity increases. However, just after January 14th, the temperature and humidity drops. 