---
title: "p8105_hw1_ksf2129"
author: Kimia Faroughi
output: github_document
date: "2025-09-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(moderndive)
library(tidyverse)
```

## Problem 1

Load the early_january_weather dataset.

```{r}
data("early_january_weather")
```

Look at the data.

```{r}
early_january_weather #view dataset
names(early_january_weather) #print variable names
help(early_january_weather) #open dataset help file
nrow(early_january_weather) #print number of rows
ncol(early_january_weather) #print number of columns
skimr::skim(early_january_weather) #skim data (to find mean of temp)
```

# Description of the Dataset

The dataset shows hourly meteorological data for LGA, JFK, and EWR airports for January 2013. 

Variables include the "origin" (weather station/airport), year, month (with 1 denoting January), day, hour (starting with 0 and going to 23), "temp" (temperature in Fahrenheit), "dewp" (dewpoint in Fahreinheit), "humid" (relative humidity), "wind_dir" (wind direction in degrees), "wind_speed" (wind speed in miles per hour), "wind_gust" (gust speed in miles per hour), "precip" (precipitation in inches), "pressure" (sea level pressure in millibars), "visib" (visibility in miles), and "time_hour" (date and hour of the recording). 

There are 358 rows (observations) and 15 columns (variables) in the dataset.

The mean temperature (in Fahrenheit) is 3.958212e+01 = 39.58212. 

# Scatterplot

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point() #generate scatterplot
ggsave("problem1_scatterplot.pdf", height = 4, width = 6) #save scatterplot
```

The scatterplot shows a wave pattern throughout the days of the month.